---
title: "sub clustering striatal clusters"
author: "Yuvarani Masarapu & Marta Graziano"
date: '2022-08-24'
output: html_document
---


```{r}
summarizeSweep <- function(sweep.list, GT = FALSE, GT.calls = NULL) {
  require(KernSmooth); require(ROCR)
  ## Set pN-pK param sweep ranges
  name.vec <- names(sweep.list)
  name.vec <- unlist(strsplit(name.vec, split="pN_"))
  name.vec <- name.vec[seq(2, length(name.vec), by=2)]
  name.vec <- unlist(strsplit(name.vec, split="_pK_"))
  pN <- as.numeric(unique(name.vec[seq(1, length(name.vec), by=2)]))
  pK <- as.numeric(unique(name.vec[seq(2, length(name.vec), by=2)]))

  ## Initialize data structure w/ or w/o AUC column, depending on whether ground-truth doublet classifications are available
  if (GT == TRUE) {
    sweep.stats <- as.data.frame(matrix(0L, nrow=length(sweep.list), ncol=4))
    colnames(sweep.stats) <- c("pN","pK","AUC","BCreal")
    sweep.stats$pN <- factor(rep(pN, each=length(pK), levels = pN))
    sweep.stats$pK <- factor(rep(pK, length(pN),levels = pK))
  }

  if (GT == FALSE) {
    sweep.stats <- as.data.frame(matrix(0L, nrow=length(sweep.list), ncol=3))
    colnames(sweep.stats) <- c("pN","pK","BCreal")
    sweep.stats$pN <- factor(rep(pN, each=length(pK), levels = pN))
    sweep.stats$pK <- factor(rep(pK, length(pN),levels = pK))
  }

  ## Perform pN-pK parameter sweep summary
  for (i in 1:length(sweep.list)) {
    res.temp <- sweep.list[[i]]

    ## Use gaussian kernel density estimation of pANN vector to compute bimodality coefficient
    gkde <- approxfun(bkde(res.temp$pANN, kernel="normal"))
    x <- seq(from=min(res.temp$pANN), to=max(res.temp$pANN), length.out=nrow(res.temp))
    sweep.stats$BCreal[i] <- bimodality_coefficient(gkde(x))

    if (GT == FALSE) { next }

    ## If ground-truth doublet classifications are available, perform ROC analysis on logistic
    ## regression model trained using pANN vector
    meta <- as.data.frame(matrix(0L, nrow=nrow(res.temp), ncol=2))
    meta[,1] <- GT.calls
    meta[,2] <- res.temp$pANN
    train.ind <- sample(1:nrow(meta), round(nrow(meta)/2), replace=FALSE)
    test.ind <- (1:nrow(meta))[-train.ind]
    colnames(meta) <- c("SinDub","pANN")
    meta$SinDub <- factor(meta$SinDub, levels = c("Doublet","Singlet"))
    model.lm <- glm(SinDub ~ pANN, family="binomial"(link='logit'), data=meta, subset=train.ind)
    prob <- predict(model.lm, newdata=meta[test.ind, ], type="response")
    ROCpred <- ROCR::prediction(predictions=prob, labels=meta$SinDub[test.ind])
    perf.auc <- ROCR::performance(ROCpred, measure="auc")
    sweep.stats$AUC[i] <- perf.auc@y.values[[1]]
  }

  return(sweep.stats)

}
parallel_paramSweep <- function(n, n.real.cells, real.cells, pK, pN, data, orig.commands, PCs, sct)  {

  sweep.res.list = list()
  list.ind = 0

  ## Make merged real-artifical data
  print(paste("Creating artificial doublets for pN = ", pN[n]*100,"%",sep=""))
  n_doublets <- round(n.real.cells/(1 - pN[n]) - n.real.cells)
  real.cells1 <- sample(real.cells, n_doublets, replace = TRUE)
  real.cells2 <- sample(real.cells, n_doublets, replace = TRUE)
  doublets <- (data[, real.cells1] + data[, real.cells2])/2
  colnames(doublets) <- paste("X", 1:n_doublets, sep = "")
  data_wdoublets <- cbind(data, doublets)

  ## Pre-process Seurat object
  if (sct == FALSE) {
    print("Creating Seurat object...")
    seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

    print("Normalizing Seurat object...")
    seu_wdoublets <- NormalizeData(seu_wdoublets,
                                   normalization.method = orig.commands$NormalizeData.RNA@params$normalization.method,
                                   scale.factor = orig.commands$NormalizeData.RNA@params$scale.factor,
                                   margin = orig.commands$NormalizeData.RNA@params$margin)

    print("Finding variable genes...")
    seu_wdoublets <- FindVariableFeatures(seu_wdoublets,
                                          selection.method = orig.commands$FindVariableFeatures.RNA$selection.method,
                                          loess.span = orig.commands$FindVariableFeatures.RNA$loess.span,
                                          clip.max = orig.commands$FindVariableFeatures.RNA$clip.max,
                                          mean.function = orig.commands$FindVariableFeatures.RNA$mean.function,
                                          dispersion.function = orig.commands$FindVariableFeatures.RNA$dispersion.function,
                                          num.bin = orig.commands$FindVariableFeatures.RNA$num.bin,
                                          binning.method = orig.commands$FindVariableFeatures.RNA$binning.method,
                                          nfeatures = orig.commands$FindVariableFeatures.RNA$nfeatures,
                                          mean.cutoff = orig.commands$FindVariableFeatures.RNA$mean.cutoff,
                                          dispersion.cutoff = orig.commands$FindVariableFeatures.RNA$dispersion.cutoff)

    print("Scaling data...")
    seu_wdoublets <- ScaleData(seu_wdoublets,
                               features = orig.commands$ScaleData.RNA$features,
                               model.use = orig.commands$ScaleData.RNA$model.use,
                               do.scale = orig.commands$ScaleData.RNA$do.scale,
                               do.center = orig.commands$ScaleData.RNA$do.center,
                               scale.max = orig.commands$ScaleData.RNA$scale.max,
                               block.size = orig.commands$ScaleData.RNA$block.size,
                               min.cells.to.block = orig.commands$ScaleData.RNA$min.cells.to.block)

    print("Running PCA...")
    seu_wdoublets <- RunPCA(seu_wdoublets,
                            features = orig.commands$ScaleData.RNA$features,
                            npcs = length(PCs),
                            rev.pca =  orig.commands$RunPCA.RNA$rev.pca,
                            weight.by.var = orig.commands$RunPCA.RNA$weight.by.var,
                            verbose=FALSE)
  }

  if (sct == TRUE) {
    require(sctransform)
    print("Creating Seurat object...")
    seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

    print("Running SCTransform...")
    seu_wdoublets <- SCTransform(seu_wdoublets)

    print("Running PCA...")
    seu_wdoublets <- RunPCA(seu_wdoublets, npcs = length(PCs))
  }

  ## Compute PC distance matrix
  print("Calculating PC distance matrix...")
  nCells <- nrow(seu_wdoublets@meta.data)
  pca.coord <- seu_wdoublets@reductions$pca@cell.embeddings[ , PCs]
  rm(seu_wdoublets)
  gc()
  dist.mat <- fields::rdist(pca.coord)[,1:n.real.cells]

  ## Pre-order PC distance matrix prior to iterating across pK for pANN computations
  print("Defining neighborhoods...")
  for (i in 1:n.real.cells) {
    dist.mat[,i] <- order(dist.mat[,i])
  }

  ## Trim PC distance matrix for faster manipulations
  ind <- round(nCells * max(pK))+5
  dist.mat <- dist.mat[1:ind, ]

  ## Compute pANN across pK sweep
  print("Computing pANN across all pK...")
  for (k in 1:length(pK)) {
    print(paste("pK = ", pK[k], "...", sep = ""))
    pk.temp <- round(nCells * pK[k])
    pANN <- as.data.frame(matrix(0L, nrow = n.real.cells, ncol = 1))
    colnames(pANN) <- "pANN"
    rownames(pANN) <- real.cells
    list.ind <- list.ind + 1

    for (i in 1:n.real.cells) {
      neighbors <- dist.mat[2:(pk.temp + 1),i]
      pANN$pANN[i] <- length(which(neighbors > n.real.cells))/pk.temp
    }

    sweep.res.list[[list.ind]] <- pANN

  }

  return(sweep.res.list)
}
paramSweep <- function(seu, PCs=1:10, sct = FALSE, num.cores=1) {
  require(Seurat); require(fields); require(parallel)
  ## Set pN-pK param sweep ranges
  pK <- c(0.0005, 0.001, 0.005, seq(0.01,0.3,by=0.01))
  pN <- seq(0.05,0.3,by=0.05)

  ## Remove pK values with too few cells
  min.cells <- round(nrow(seu@meta.data)/(1-0.05) - nrow(seu@meta.data))
  pK.test <- round(pK*min.cells)
  pK <- pK[which(pK.test >= 1)]

  ## Extract pre-processing parameters from original data analysis workflow
  orig.commands <- seu@commands

  ## Down-sample cells to 10000 (when applicable) for computational effiency
  if (nrow(seu@meta.data) > 10000) {
    real.cells <- rownames(seu@meta.data)[sample(1:nrow(seu@meta.data), 10000, replace=FALSE)]
    data <- seu@assays$RNA$counts[ , real.cells]
    n.real.cells <- ncol(data)
  }

  if (nrow(seu@meta.data) <= 10000){
    real.cells <- rownames(seu@meta.data)
    data <- seu@assays$RNA$counts
    n.real.cells <- ncol(data)
  }

  ## Iterate through pN, computing pANN vectors at varying pK
  #no_cores <- detectCores()-1
  if(num.cores>1){
    require(parallel)
    cl <- makeCluster(num.cores)
    output2 <- mclapply(as.list(1:length(pN)),
                        FUN = parallel_paramSweep,
                        n.real.cells,
                        real.cells,
                        pK,
                        pN,
                        data,
                        orig.commands,
                        PCs,
                        sct,mc.cores=num.cores)
    stopCluster(cl)
  }else{
    output2 <- lapply(as.list(1:length(pN)),
                      FUN = parallel_paramSweep,
                      n.real.cells,
                      real.cells,
                      pK,
                      pN,
                      data,
                      orig.commands,
                      PCs,
                      sct)
  }

  ## Write parallelized output into list
  sweep.res.list <- list()
  list.ind <- 0
  for(i in 1:length(output2)){
    for(j in 1:length(output2[[i]])){
      list.ind <- list.ind + 1
      sweep.res.list[[list.ind]] <- output2[[i]][[j]]
    }
  }

  ## Assign names to list of results
  name.vec <- NULL
  for (j in 1:length(pN)) {
    name.vec <- c(name.vec, paste("pN", pN[j], "pK", pK, sep = "_" ))
  }
  names(sweep.res.list) <- name.vec
  return(sweep.res.list)

}
run_doublet_finder <- function(x, do_plot = T, header = "") {
    # Given the number of cells in a library, return the estimated doublet rate. 
    # Number taken from
    # https://kb.10xgenomics.com/hc/en-us/articles/360001378811-What-is-the-maximum-number-of-cells-that-can-be-profiled-
    
    doublet_rate <- function(n_cells) {
        if (n_cells < 750) {
            return(0.004)
        }
        if (n_cells < 1500) {
            return(0.008)
        }
        if (n_cells < 2500) {
            return(0.016)
        }
        if (n_cells < 3500) {
            return(0.024)
        }
        if (n_cells < 4500) {
            return(0.032)
        }
        if (n_cells < 5500) {
            return(0.04)
        }
        if (n_cells < 6500) {
            return(0.048)
        }
        if (n_cells < 7500) {
            return(0.056)
        }
        if (n_cells < 8500) {
            return(0.064)
        }
        if (n_cells < 9500) {
            return(0.072)
        }
        if (n_cells < 10500) {
            return(0.08)
        }

        # extrapolating 0.01 for each additional 1000 cells.
        if (n_cells < 11500) {
            return(0.09)
        }
        if (n_cells < 12500) {
            return(0.1)
        }
        if (n_cells < 13500) {
            return(0.11)
        }
        if (n_cells < 13500) {
            return(0.12)
        }
        if (n_cells < 14500) {
            return(0.13)
        }
        if (n_cells < 15500) {
            return(0.14)
        }
        if (n_cells < 16500) {
            return(0.16)
        } 
        else (return(0.17))
    }

    # Normalize for doublet finder
    DefaultAssay(x) <- "RNA"
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 1000)
    x <- ScaleData(x)
    x <- RunPCA(x,npcs=20, verbose = FALSE)
    x <- RunUMAP(x, dims = 1:20)

    # pK Identification (no ground-truth)
    sweep_list <- paramSweep(x, PCs = 1:10)
    sweep_stats <- summarizeSweep(sweep_list)
    bcmvny <- find.pK(sweep_stats)
    use_pK <- as.numeric(levels(bcmvny$pK)[which(bcmvny$BCmetric == max(bcmvny$BCmetric))])

    # Estimate doublet rate and predict doublets
    n_cells <- nrow(x@meta.data)
    nExp_poi <- round(doublet_rate(n_cells) * n_cells)
    x <- doubletFinder(x, PCs = 1:10, pN = 0.25, pK = use_pK, nExp = nExp_poi)

    DF.name = colnames(x@meta.data)[grepl("DF.classification", colnames(x@meta.data))]

    if (do_plot) {
        print(DimPlot(x, group.by = DF.name) + ggtitle(header))
        print(VlnPlot(x, features = "nFeature_RNA", group.by = DF.name, pt.size = 0))
    }

    out_pred <- x@meta.data[, DF.name]
    #names(out_pred) <- colnames(x)
    return(x)
    #return(out_pred)
}
```

#read count matrix in R 

```{r}
# Read count matrix into R as seurat object

#Count data from all mice samples of mitopark and 6OHDA models are combined. The sample labels are replaced to new from old labels as below.
theme_set(theme_cowplot())
setwd("Z:/dmclab/Marta/PD/snRNA human putamen/data")
indir <- "Z:/dmclab/Marta/PD/snRNA human putamen/data/cellranger_output"
data_dir <- paste(indir, "/", sep = "")
files <- dir(data_dir)

files

seurat_list <- lapply(files, function(file_n){
  print(file_n)
  
  matrix_dir <- paste(data_dir, "/", file_n, "/outs/filtered_feature_bc_matrix/", sep = "")
  
  barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
  features.path <- paste0(matrix_dir, "features.tsv.gz")
  matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
  mat <- readMM(file = matrix.path)
  
  feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
  barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
  colnames(mat) = barcode.names$V1
  
  # Ensure unique row names
  if (any(duplicated(feature.names$V2))) {
    feature.names$V2 <- make.unique(feature.names$V2, sep = "_") # Appending a suffix to make row names unique
  }
  rownames(mat) <- feature.names$V2 #we need gene names in form of ID, so we take column 2 where gene ids are
  
  # create seurat object
  cur_seurat <- CreateSeuratObject(
    counts = mat,
    project = "HumanPutamen_PD_ctrl_dataset"
  )
  cur_seurat@meta.data$SampleID <- file_n
  
  cur_seurat

})



```

```{r merge_data, warning=FALSE, message=FALSE, eval=FALSE}
# merge into one big seurat object
seurat_rna <- merge(x=seurat_list[[1]], y=seurat_list[2:length(seurat_list)])
rm(seurat_list)
```

```{r cellcyclescoring, message=FALSE, verbose=FALSE, eval=FALSE}
# Classify each cell by cell-cycle phase (G2M/G1/S) and give cellcycle scoring

#CHECK SAMPLE NAMES
IDS2<-unique(seurat_rna$SampleID)

print(IDS2)

library(stringr)
indir <- getwd()

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
# Format cell cycle genes from Seurat, e.g. convert 'MCM5' to 'Mcm5'
s.genes <- cc.genes$s.genes %>%
              str_to_upper
g2m.genes <- cc.genes$g2m.genes %>%
              str_to_upper 
all.genes <- rownames(seurat_rna)

g2m.genes <- intersect(g2m.genes, all.genes)
s.genes <- intersect(s.genes, all.genes)

seurat_rna <- CellCycleScoring(object = seurat_rna, assay = "RNA", g2m.features = g2m.genes, s.features = s.genes)

seurat_rna$Sample_group <- colnames(seurat_rna)

seurat_rna$Sample_group[seurat_rna$SampleID %in% c( "AA_ASAP117_PD_NP22-55_PUT", "AA_ASAP138_PD_NP16-285_PUT","AA_ASAP139_PD_NP17-232_PUT","AA_ASAP140_PD_NP19-91_PUT", "AA_ASAP144_PD_NP19-137_PUT",  "AA_ASAP146_PD_NP21-04_PUT","AA_ASAP149_PD_NP23-21_PUT", "AA_ASAP87_PD_NP21-208_PUT" , "AA_ASAP88_PD_NP21-217_PUT","AA_ASAP97_PD_NP21-57_PUT", "AA71_ASAP70_PD_NP18-304_PUT","AA73_ASAP65_PD_NP16-269_PUT",   "AA74_ASAP62_PD_NP16-140_PUT" ,  "AA76_ASAP74_PD_NP19-255_PUT" ,"AA77_ASAP61_PD_NP16-25_PUT","AA78_ASAP63_PD_NP16-160_PUT","AA79_ASAP64_PD_NP16-162_PUT","AA80_ASAP66_PD_NP17-94_PUT",  "AA81_ASAP67_PD_NP17-191_PUT", "AA82_ASAP68_PD_NP18-117_PUT","AA83_ASAP69_PD_NP18-287_PUT", "AA84_ASAP71_PD_NP19-16_PUT" , "AA85_ASAP72_PD_NP19-23_PUT","AA86_ASAP73_PD_NP19-108_PUT","Seq176_1", "Seq176_10"   ,"Seq176_11" ,"Seq176_12","Seq176_13" , "Seq176_14","Seq176_2","Seq176_3","Seq176_4","Seq176_5","Seq176_6" , "Seq176_7","Seq176_8", "Seq176_9"  )] <- "PD"
seurat_rna$Sample_group[seurat_rna$SampleID %in% c("AA_ASAP100_ctrl_NP18-148_PUT","AA_ASAP141_ctrl_NP19-218_PUT","AA_ASAP142_ctrl_NP22-37_PUT","AA_ASAP143_ctrl_NP22-75_PUT", "AA_ASAP92_ctrl_NP16-284_PUT",  "AA_ASAP93_ctrl_NP16-293_PUT", "AA_ASAP94_ctrl_NP17-20_PUT" ,"AA_ASAP95_ctrl_NP17-216_PUT","AA_ASAP96_ctrl_NP18-46_PUT","AA_ASAP98_ctrl_NP16-119_PUT" ,"AA87_ASAP76_ctrl_NP18-159_PUT", "AA88_ASAP77_ctrl_NP19-36_PUT","AA89_ASAP78_ctrl_NP19-37_PUT" ,"Seq176_15", "Seq176_16","Seq176_17" )] <- "Control"

                                                 

saveRDS(seurat_rna, file = "Z:/dmclab/Marta/PD/snRNA human putamen/data/raw_combined_snRNAseq_full_dataset.rds")
#seurat_rna<- readRDS("Z:/dmclab/Marta/PD/snRNA human putamen/data/raw_combined_snRNAseq.rds")

# use table function to get the number of cells in each Sample as a dataframe
df <- as.data.frame(rev(table(seurat_rna$SampleID)))
colnames(df) <- c('Sample', 'n_cells')

# bar plot of the number of cells in each sample
p <- ggplot(df, aes(y=n_cells, x=reorder(Sample, Sample), fill=Sample)) +
  geom_bar(stat='identity') +
  geom_text(aes(label = n_cells), vjust = 1.5, colour = "black") +
  scale_y_continuous(expand = c(0,0)) +
  NoLegend() + RotatedAxis() +
  ylab(expression(italic(N)[cells])) + xlab('Sample Name') +
  ggtitle(paste('Total cells:', sum(df$n_cells))) +
  theme(
    panel.grid.minor=element_blank(),
    panel.grid.major.y=element_line(colour="lightgray", size=0.5),
  )

#png('figures/basic_cells_per_sample.png', width=9, height=4, res=200, units='in')
quartz()
p
```

```{r}
# Plot most abundant transctipts
par(mar = c(4, 8, 2, 1))
C <- seurat_rna@assays$RNA@counts
C <- Matrix::t(Matrix::t(C)/Matrix::colSums(C)) * 100
saveRDS(C, file = "C.rds")
most_expressed <- order(apply(C, 1, median), decreasing = T)[20:1]
saveRDS(most_expressed, file = "most_expressed.rds")
# Plot most abundant transctipts
par(mar = c(4, 8, 2, 1))
C <- readRDS("C.rds")
most_expressed <- readRDS("most_expressed.rds")
boxplot(as.matrix(t(C[most_expressed, ])), cex = 0.1, las = 1, xlab = "% total count per cell", col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)

quartz()
p
```


```{r}
#violin plot with raw values
p1 <- VlnPlot(object = seurat_rna, features = c("nFeature_RNA"), group.by = "SampleID",
    pt.size = 0, log = F)

p2 <- VlnPlot(object = seurat_rna, features = c("nCount_RNA"), group.by = "SampleID",
    pt.size = 0, log = F)
quartz()
cowplot::plot_grid(plotlist = list(p1,p2), nrow = 1, ncol = 2)

#violin plot with logscaled values
p1 <- VlnPlot(object = seurat_rna, features = c("nFeature_RNA"), group.by = "SampleID",
    pt.size = 0, log = T)

p2 <- VlnPlot(object = seurat_rna, features = c("nCount_RNA"), group.by = "SampleID",
    pt.size = 0, log = T)
quartz()
cowplot::plot_grid(plotlist = list(p1,p2), nrow = 1, ncol = 2)

rm(p1,p2)
```

After filtering
Overexpressed Malat1 gene,
mitochondrial and ribosomal genes
All genes with 0 transcripts detected
nFeature_RNA > 300 & nFeature_RNA < 11000 & nCount_RNA > 350 &  nCount_RNA < 75000
Genes per UMI per cell ratio above 0.82 (gives an idea of the complexity of each cell)

```{r}

#remove counts variable from previous step
rm(C)
# Define the genes to be removed

# Filter out genes from the Seurat object
# Define the genes to be removed
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("biomaRt")
library(biomaRt)

mart <- useMart(biomart="ensembl", dataset="hsapiens_gene_ensembl")
Y_Genes <- getBM(attributes = c("chromosome_name",  "hgnc_symbol"),
           filters = "chromosome_name", values = "Y", mart = mart)
X_Genes <- getBM(attributes = c("chromosome_name",  "hgnc_symbol"),
           filters = "chromosome_name", values = "X", mart = mart)

genes_to_remove <- c("MALAT1", "XIST", "UTY", "TSIX", "DDX3Y", "DDX3X", "TBL1X", "USP9X", "JPX", "FTX", "TSNAX", "PCDH11X", "EIF2S3Y", "EIF2S3X", "TMSB4X", "TMSB4Y", "KDM5D")
genes_to_remove<- c(genes_to_remove,Y_Genes$hgnc_symbol, X_Genes$hgnc_symbol )
# Filter out genes from the Seurat object
seurat_rna <- seurat_rna[!rownames(seurat_rna) %in% genes_to_remove, ]
#seurat_rna = JoinLayers(seurat_rna)
# Remove mitochondrial genes, 13 genes filtered out
seurat_rna[["percent_mt"]] <- PercentageFeatureSet(seurat_rna, pattern = "^MT-")
seurat_rna <- seurat_rna[!grepl("^MT-", rownames(seurat_rna)), ]

# Remove ribosomal genes, 111 genes filtered out
seurat_rna <- seurat_rna[!grepl("^RPS|^RPL", rownames(seurat_rna)), ]

# log10 of number of genes detected per UMI
seurat_rna$log10GenesPerUMI <- log10(seurat_rna$nFeature_RNA) / log10(seurat_rna$nCount_RNA)
seurat_rna$orig.ident<- "snRNA seq"
library(data.table)
df <- as.data.table(seurat_rna@meta.data)
sel <- c("orig.ident", "nCount_RNA", "nFeature_RNA", "percent_mt", "log10GenesPerUMI")
df <- df[, sel, with = FALSE]
#df[1:3, ]


fontsize <- 10
linesize <- 0.35

gp.ls <- df[, 2:5] %>% imap( ~ {
  
   # define lable fun
  give.n <- function(x) {
    return(c(y = median(x) + max(x) / 10, label = round(median(x), 2)))
  }
  
  # assign colors
  col.ls <-
    setNames(
      c('lightpink2', 'lightblue2', 'lightgreen', 'coral1'),
      c("nCount_RNA", "nFeature_RNA", "percent_mt", "log10GenesPerUMI")
    )
  
  ggplot(data = df, aes(x = orig.ident, y = .x)) +
    geom_violin(trim = FALSE, fill = col.ls[.y]) +
    ggtitle(label = .y) + ylab(label = .y) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      strip.background = element_blank(),
      panel.border = element_blank()
    ) +
    theme(
      axis.text = element_text(size = fontsize),
      axis.line = element_line(colour = "black", size = linesize),
      axis.ticks = element_line(size = linesize),
      axis.title.x = element_blank(),
      axis.ticks.length = unit(.05, "cm"),
      plot.title = element_text(size = fontsize + 2, hjust = 0.5),
      legend.position = 'none'
    ) +
    stat_summary(fun = median, geom = "point", col = "black") +  # Add points to plot
    stat_summary(fun.data = give.n,
                 geom = "text",
                 col = "black")
})

 grid.arrange(gp.ls[[1]], gp.ls[[2]], gp.ls[[3]], gp.ls[[4]], ncol = 2)
# Add metadata back to Seurat object
metadata <- seurat_rna@meta.data

# Visualize the number UMIs/transcripts per cell
metadata %>% 
  	ggplot(aes(color=SampleID, x=nCount_RNA, fill= SampleID)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  labs(title="Visualize the number UMIs/transcripts per cell")+
  	geom_vline(xintercept = c(400, 75000))

# Visualize the number genes per cell
metadata %>% 
  	ggplot(aes(color=SampleID, x=nFeature_RNA, fill= SampleID)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  labs(title=" Visualize the number genes per cell")+
  	geom_vline(xintercept = c(300, 10000))

# Visualize the number genes per cell
metadata %>% 
  	ggplot(aes(color=SampleID, x=log10GenesPerUMI, fill= SampleID)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = c(0.82))
```

```{r subsetting, warning=FALSE}
# subset will do cell-level filtering, remove cells with detected genes less than 350
seurat_rna <- subsetseurat_rna <- subset(seurat_rna, subset = nFeature_RNA > 300 & nFeature_RNA < 11000 & nCount_RNA > 350 &  nCount_RNA < 75000)


# complexity of each cell to be > 0.80, so genes detected proportional to UMI
seurat_rna <- subset(seurat_rna, subset = log10GenesPerUMI > 0.8)

# Always do gene level threshold filtering AFTER cell filtering. 
# There could be some rows or genes with '0' transcripts that can remain after the cell filtering and if done before, this may lead to some rows in the matrix with all '0' values across.
# We also want to remove genes with 0 transcripts. Using subset() will not work here
seurat_rna <- seurat_rna[rowSums(seurat_rna) > 0, ]

```
Check data quality after filtering
```{r}
p1 <- VlnPlot(object = seurat_rna, features = c("nFeature_RNA"), group.by = "Sample",
    pt.size = 0, log = F)

p2 <- VlnPlot(object = seurat_rna, features = c("nCount_RNA"), group.by = "Sample",
    pt.size = 0, log = F)
quartz()
cowplot::plot_grid(plotlist = list(p1,p2), nrow = 1, ncol = 2)
#y axis with log scale

p1 <- VlnPlot(object = seurat_rna, features = c("nFeature_RNA"), group.by = "Sample",
    pt.size = 0, log = T)

p2 <- VlnPlot(object = seurat_rna, features = c("nCount_RNA"), group.by = "Sample",
    pt.size = 0, log = T)

cowplot::plot_grid(plotlist = list(p1,p2), nrow = 1, ncol = 2)

#Percentage mitchondria per nuclei-Removed all mt- genes but let’s see how they were distributed across each cell/nuclei

VlnPlot(object = seurat_rna, features = c("percent_mt"), group.by = "SampleID",
    pt.size = 0, log = T)

```

```{r}


data.filt <- SplitObject(seurat_rna, split.by = "SampleID")


#Main clusters object
data.object.filter.norm.integrated <- "Z:/dmclab/Marta/PD/snRNA human putamen/data/rna.combined.snRNA_seq_merged_42_res0.4_2_annotated.rds"


```

We have sample numbers added at the end of each spot id with an '_' in the main clusters object.
But for the original raw object list, this is not the case, hence we need to first check if the order of the samples is the same for both objects.

```{r}
for(id in 1:length(data.filt)){
  print(unique(data.filt[[id]]$SampleID))
}
```

```{r}
ids<- unique(data.object.filter.norm.integrated$SampleID)

# Subset the Seurat object based on those IDs
data.filt1 <- subset(x = data.filt, cells = WhichCells(data.filt, expression = SampleID %in% ids))


for(id in 1:length(data.filt)){
  print(unique(data.filt[[id]]$SampleID))
}

```

If yes, then we go ahead and merge the filtered object list, add sample numbers to each spot id. This way, the spot ids will match and we will be able to filter the spots we want to keep for the striatal sub-clustering.

```{r}
#Here, seurat will automatically add sample numbers suffixed to each spot id to make sure they are unique.

temp <- merge(data.filt[[1]], y = data.filt[2:length(data.filt)], project = "snRNA seq")

#sanity check to make sure that the sample order is the same as in striatal_clus object
unique(temp$SampleID)

#sanity check to make sure that the spot ids follow the same pattern as in the striatal_clus object i.e <spot_id>_<sample_no.>
head(colnames(temp))
```

We filter the merged object to keep only the spots from striatal clusters for the sub-clustering.

```{r}
unique(data.object.filter.norm.integrated@active.ident)
#data.object.filter.norm.integrated[["RNA"]] <-JoinLayers(data.object.filter.norm.integrated[["RNA"]])

cells.keep <- colnames(data.object.filter.norm.integrated)[data.object.filter.norm.integrated@active.ident %in% c("eSPNs","SPNs")]

head(cells.keep)
```

```{r}
temp <- temp[ ,colnames(temp) %in% cells.keep]

#double check as a sanity check
unique(colnames(temp) == cells.keep)
length(colnames(temp)) == length(cells.keep)
```

Now we unlist the object again in order to perform SCTransform by sample. Then do downstream analysis as we would normally run.

```{r}
data.filt <- SplitObject(object = temp, split.by = "SampleID")

#change list names from sample ids to numbers
print(names(data.filt) <- c(1:50))

saveRDS(data.filt, file = "Z:/dmclab/Marta/PD/snRNA human putamen/data/parkinsons_brain_sub-clusters_filt_allen_annotation.rds")
```

Check again if the sample order matches the data.object.filter.norm.integrated object

```{r}
for(id in 1:length(data.filt)){
  print(unique(data.filt[[id]]$SampleID))
  head(colnames(data.filt[[1]]))
}

#Here the correct sample numbers should show as suffixes to spot ids
for(id in 1:length(data.filt)){
  print(head(colnames(data.filt[[id]])))
}

```


```{r}
#seurat_rna_list <- SplitObject(data.filt, split.by = "Sample")

#names(seurat_rna_list)
set.seed(1)

all_doublets <- NULL
all_doublet_scores <- NULL
pANN_max <- 1

for (x_name in names(data.filt)){

    x <- data.filt[[x_name]]

    VlnPlot(x, features = c("nFeature_RNA", "nCount_RNA", "percent_mt"), ncol = 3,
        log = T, pt.size = 0) + NoLegend()

    x <- run_doublet_finder(x, do_plot = T, header = x_name)
    doublet_cutoff <- 0.6
    pANN.name = colnames(x@meta.data)[grepl("pANN", colnames(x@meta.data))]
    DF.name = colnames(x@meta.data)[grepl("DF", colnames(x@meta.data))]
    new_doublet_scores <- x[[pANN.name]][, 1]
    names(new_doublet_scores) <- rownames(x[[pANN.name]])
    new_doublet_calls <- x[[DF.name]][, 1]
    names(new_doublet_calls) <- rownames(x[[DF.name]])

    #final cells which are labelled doublets
    new_doublets <- union(names(which(new_doublet_calls == "Doublet")), names(which(new_doublet_scores >=
        pANN_max)))
    x <- x[, setdiff(colnames(x), new_doublets)]
    all_doublets <- union(all_doublets, new_doublets)
    all_doublet_scores <- c(all_doublet_scores, new_doublet_scores)

    #filtering step again to remove rows with '0' transcripts
    x <- x[rowSums(x) > 0, ]

    # adding vst.flavour = "v2" makes values 0 across rows for some genes.
    x <- SCTransform(x, verbose = T, variable.features.n = 6000)

    data.filt[[x_name]] <- x
}


rm(x)

for(i in 1:length(data.filt)){
  data.filt[[i]] <- AddMetaData(
  object = data.filt[[i]] ,
  metadata = all_doublet_scores[colnames(data.filt[[i]])],
  col.name = "pANN"
)
}


# saveRDS(data.object.filter.norm, file = "Z:/dmclab/Marta/snRNAseq_mitopark/data/parkinsons_brain_sub-clusters_norm.rds")


```

Downstream processing

```{r eval=FALSE}
# select features that are repeatedly variable across datasets for integration
#CHECK DIFFERENT nfEATURES AND CHECK WHEN IT STOPS AND THEN TAKE THE TOP 50% VARIABLE GENES OR WHATEVER PERCENTAGE YOU WANT:
#Max variable features foud: 9242


# Assuming data.filt is your list of filtered Seurat objects

#filtering step in which i remove objects that do not pass the filtering criteria, which is minimum number of cells: elements with a minimum number of columns. 
data.filt1 <- Filter(function(x) ncol(x@assays$RNA@counts) > 30, data.filt)

for (i in seq_along(data.filt1)){
  
  x <- data.filt1[[i]]
  DefaultAssay(x) <- "RNA"
  x <- SCTransform(x, verbose = FALSE)
  data.filt1[[i]] <- x

  
}

# Check if SCT models are present
for (i in seq_along(data.filt1)) {
  if (!"SCT" %in% Assays(data.filt1[[i]])) {
    stop(paste("SCT model not found in Seurat object at index", i))
  }
}

features <- SelectIntegrationFeatures(object.list = data.filt1, nfeatures = 6000)
data.filt1 <- PrepSCTIntegration(object.list = data.filt1, anchor.features = features)

# Step 5: Determine the appropriate number of dimensions for anchor finding
min_cells <- min(sapply(data.filt1, function(x) ncol(x@assays$RNA@counts)))
max_dims <- min(30, min_cells - 1)

# Print diagnostic information
cat("Minimum number of cells:", min_cells, "\n")
cat("Max dimensions for anchor finding:", max_dims, "\n")

max_ncol <- 0
max_seurat <- NULL
for (i in seq_along(data.filt1)) {
    seurat_rna <- data.filt1[[i]]  # Get the Seurat object at index i
    
    # Get the number of columns for the current Seurat object
    current_ncol <- ncol(seurat_rna)
    
    # Check if the number of columns for the current Seurat object is greater than the maximum number of columns encountered so far
    if (current_ncol > max_ncol) {
        # Update the maximum number of columns and the corresponding Seurat object
        max_ncol <- current_ncol
        max_seurat <- seurat_rna
        max_index <- i  # Store the index of the maximum Seurat object
    }
}

# Step 6: Find integration anchors
rna.anchors <- FindIntegrationAnchors(object.list = data.filt1, normalization.method = "SCT", anchor.features = features, reference =max_index)
#use the max index in the function
saveRDS(rna.anchors, file = "Z:/dmclab/Marta/Combined_snRNAseq/data/rna.anchors_subclustering.rds")
rna.combined <- IntegrateData(anchorset = rna.anchors, k.weight=30)
rna.combined <- RunPCA(object = rna.combined , assay = "SCT", features = features, npcs = 50, verbose = FALSE, reduction.name = "pca_before_harmony", seed.use = 17)
saveRDS(rna.combined, file = "Z:/dmclab/Marta/PD/snRNA human putamen/data/rna.combined_subclustering_allen_1206.rds")
#rna.combined <- readRDS(file = "rna.combined.clusters.rds")
e<-ElbowPlot(object = rna.combined, ndims = 50, reduction = "pca_before_harmony") + ggtitle("Elbow plot to select significant PCs")

```

```{r}
rna.combined <- RunUMAP(object = rna.combined, dims = 1:42, assay = "SCT", seed.use = 63, reduction = "pca_before_harmony", reduction.name = "umap_before_harmony") #36 contributing PCs chosen for further analysis

p1 <- DimPlot(object = rna.combined, group.by = "SampleID", reduction = "umap_before_harmony") + ggtitle("Before batch correction (spots grouped by sample)")+ theme(legend.position = "none")

p2 <- DimPlot(object = rna.combined, group.by = "Sample_group",reduction = "umap_before_harmony")  + ggtitle("sample groups")+ theme(legend.position = "none")

p3 <- FeaturePlot(object = rna.combined, features = "nCount_SCT", reduction = "umap_before_harmony")
p4 <- FeaturePlot(object = rna.combined, features = "nCount_RNA", reduction = "umap_before_harmony")

p5<- (p1 - p2) / (p3 - p4)
plot_directory<-"Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering"

  plot_filename <- paste0(plot_directory, "/Before batch correction", ".pdf")
  pdf(plot_filename, width = 18, height = 18)  # Adjust width and height as desired
  print(p5)
  dev.off()
```

```{r eval=FALSE}
library(harmony)
set.seed(57)
rna.combined <- RunHarmony(object = rna.combined, group.by.vars = c("SampleID"), theta = c(1), assay.use = "SCT", reduction = "pca_before_harmony", dims.use = 1:42, plot_convergence = T, verbose = F, reduction.save = "harmony_sid_1")

rna.combined <- RunUMAP(object = rna.combined, assay.use = "SCT", reduction = "harmony_sid_1", dims = 1:42, seed.use = 6129, reduction.name = "umap_after_harmony")

set.seed(57)
rna.combined <- RunHarmony(object = rna.combined, group.by.vars = c("Sample_group"), theta = c(1), assay.use = "SCT", reduction = "pca_before_harmony", dims.use = 1:42, plot_convergence = T, verbose = F, reduction.save = "harmony_sid_2")

rna.combined <- RunUMAP(object = rna.combined, assay.use = "SCT", reduction = "harmony_sid_2", dims = 1:42, seed.use = 6129, reduction.name = "umap_after_harmony_2")

set.seed(57)
rna.combined <- RunHarmony(object = rna.combined, group.by.vars = c("SampleID","Sample_group"), theta = c(1,1), assay.use = "SCT", reduction = "pca_before_harmony", dims.use = 1:42, plot_convergence = T, verbose = F, reduction.save = "harmony_sid_3")

rna.combined <- RunUMAP(object = rna.combined, assay.use = "SCT", reduction = "harmony_sid_3", dims = 1:42, seed.use = 6129, reduction.name = "umap_after_harmony_3")

set.seed(57)
rna.combined <- RunHarmony(object = rna.combined, group.by.vars = c("SampleID","Sample_group"), theta = c(2,1), assay.use = "SCT", reduction = "pca_before_harmony", dims.use = 1:42, plot_convergence = T, verbose = F, reduction.save = "harmony_sid_4")

rna.combined <- RunUMAP(object = rna.combined, assay.use = "SCT", reduction = "harmony_sid_4", dims = 1:42, seed.use = 6129, reduction.name = "umap_after_harmony_4")


```

```{r after_batcheffects, fig.width=16, fig.height=14, fig.show='hold'}
p1 <- DimPlot(object = rna.combined, group.by = "SampleID", reduction = "umap_after_harmony") + ggtitle("After batch correction (spots grouped by sample)")+ theme(legend.position = "none")

p2 <- DimPlot(object = rna.combined, group.by = "Sample_group", reduction = "umap_after_harmony",  cols=c("red","blue"))  + ggtitle("sample groups")
p3 <- FeaturePlot(object = rna.combined, features = "nCount_SCT", reduction = "umap_after_harmony")
p4 <- FeaturePlot(object = rna.combined, features = "nCount_RNA", reduction = "umap_after_harmony")


p5<- (p1 - p2) / (p3 - p4)
  plot_filename <- paste0(plot_directory, "/After batch correction (spots grouped by sample)", ".pdf")
  pdf(plot_filename, width = 18, height = 18)  # Adjust width and height as desired
  print(p5)
  dev.off()
```

```{r after_batcheffects, fig.width=16, fig.height=14, fig.show='hold'}
p1 <- DimPlot(object = rna.combined, group.by = "SampleID", reduction = "umap_after_harmony_2") + ggtitle("After batch correction (spots grouped by sample group)")+ theme(legend.position = "none")

p2 <- DimPlot(object = rna.combined, group.by = "Sample_group", reduction = "umap_after_harmony_2", cols=c( "red","blue"))  + ggtitle("sample groups")

p3 <- FeaturePlot(object = rna.combined, features = "nCount_SCT", reduction = "umap_after_harmony_2")
p4 <- FeaturePlot(object = rna.combined, features = "nCount_RNA", reduction = "umap_after_harmony_2")


p5<- (p1 - p2) / (p3 - p4)
  plot_filename <- paste0(plot_directory, "/After batch correction (spots grouped by  sample group)", ".pdf")
  pdf(plot_filename, width = 18, height = 18)  # Adjust width and height as desired
  print(p5)
  dev.off()
```

```{r after_batcheffects, fig.width=16, fig.height=14, fig.show='hold'}
p1 <- DimPlot(object = rna.combined, group.by = "SampleID", reduction = "umap_after_harmony_3") + ggtitle("After batch correction (spots grouped by sample and sample group)") + theme(legend.position="none")

p2 <- DimPlot(object = rna.combined, group.by = "Sample_group", reduction = "umap_after_harmony_3",  cols=c( "red","blue"))  + ggtitle("sample groups")

p3 <- FeaturePlot(object = rna.combined, features = "nCount_SCT", reduction = "umap_after_harmony_3")
p4 <- FeaturePlot(object = rna.combined, features = "nCount_RNA", reduction = "umap_after_harmony_3")


p5<- (p1 - p2) / (p3 - p4)
  plot_filename <- paste0(plot_directory, "/After batch correction (spots grouped by sampleID and sample group)", ".pdf")
  pdf(plot_filename, width = 18, height = 18)  # Adjust width and height as desired
  print(p5)
  dev.off()
```


```{r after_batcheffects, fig.width=16, fig.height=14, fig.show='hold'}
p1 <- DimPlot(object = rna.combined, group.by = "SampleID", reduction = "umap_after_harmony_4") + ggtitle("After batch correction (spots grouped by sample and sample group)") + theme(legend.position="none")

p2 <- DimPlot(object = rna.combined, group.by = "Sample_group", reduction = "umap_after_harmony_4",  cols=c( "red","blue"))  + ggtitle("sample groups")

p3 <- FeaturePlot(object = rna.combined, features = "nCount_SCT", reduction = "umap_after_harmony_4")
p4 <- FeaturePlot(object = rna.combined, features = "nCount_RNA", reduction = "umap_after_harmony_4")


p5<- (p1 - p2) / (p3 - p4)
  plot_filename <- paste0(plot_directory, "/After batch correction (spots grouped by sampleID THETA 2 and sample group)", ".pdf")
  pdf(plot_filename, width = 18, height = 18)  # Adjust width and height as desired
  print(p5)
  dev.off()
```

```{r}
set.seed(61)
rna.combined <- FindNeighbors(object = rna.combined, assay = "SCT", dims = 1:42, k.param = 23, graph.name = "graph_afterHarmony", reduction = "harmony_sid_4")

rna.combined <- FindClusters(object = rna.combined, pc.use = 1:42, resolution = 1, save.SNN = T, do.sparse = T, graph.name = "graph_afterHarmony", random.seed = 13, group.singletons = TRUE)
# Save the new clustering results
rna.combined$clusters_res0.6 <- rna.combined$seurat_clusters
rna.combined$clusters_res0.8 <- rna.combined$seurat_clusters

library(dplyr)
wnn_markers_RES1 <- FindAllMarkers(rna.combined, only.pos = TRUE, min.pct = 0.05, logfc.threshold = 0.05, recorrect_umi = FALSE)
wnn_markers_RES1 <- wnn_markers[wnn_markers$p_val_adj < 0.05, ]
write.csv(wnn_markers_RES1, file = "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/cluster_markers-HUMAN_snRNA_sub_allen_res1.csv", col.names = TRUE, row.names = TRUE)
saveRDS(rna.combined, file = "Z:/dmclab/Marta/PD/snRNA human putamen/data/rna.combined_subclustering_allen_1206.rds")

rna.combined<- readRDS("Z:/dmclab/Marta/PD/snRNA human putamen/data/rna.combined_subclustering_allen_1206.rds")
table1<-table("cluster" = rna.combined@active.ident, rna.combined@meta.data$SampleID)

table<- as.data.frame(table("cluster" = rna.combined_sub@active.ident, rna.combined_sub@meta.data$orig.ident))
write.csv(table, file="Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/Nuclei_per_cluster_merged.csv")


write.csv(table1, file="Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/Nuclei_across_samples.csv")
```



### Nuclei counts per sample group

```{r table3}

  # Filter by significance (adjust based on your data)
filtered_markers <- wnn_markers_RES1[wnn_markers_RES1$p_val_adj < 0.001 & wnn_markers_RES1$avg_log2FC > 0.25, ]
marker_genes <- unique(filtered_markers$gene)

library(biomaRt)


# Connect to the Ensembl BioMart database
ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# Map common gene names (HGNC symbols) to Ensembl IDs
mapping <- getBM(
  filters = "hgnc_symbol",
  attributes = c("hgnc_symbol", "ensembl_gene_id"),
  values = marker_genes,
  mart = ensembl
)

# View the mapping results
head(mapping)

# Inspect features of the Seurat object
head(rownames(rna.combined))

# Find common genes between Seurat features and your mapping
common_genes <- intersect(rownames(rna.combined), mapping$hgnc_symbol)
length(common_genes)  # Check how many genes overlap

# Subset the Seurat object based on common genes
seurat.obj <- subset(rna.combined, features = common_genes)



# Extract the normalized gene expression matrix (RNA assay)
expression_matrix <- GetAssayData(seurat.obj, assay = "RNA", slot = "data")
# Save the expression matrix
write.csv(as.matrix(expression_matrix), "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/expression_matrix_RES1.csv", row.names = TRUE)

#generate .h5ad file on python with a custom script (see .py scrpt)


#load results 

# Step 1: Load the MapMyCells results
mapmycells_results <- read.csv("Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/MAPMYCELL2/new_adata_for_MapMyCells_RES1_10xWholeHumanBrain.csv", row.names = 1)

# Step 2: Check alignment of cell names
# Inspect Seurat object cell names
head(colnames(rna.combined))

# Inspect MapMyCells result cell names
head(rownames(mapmycells_results))

# Ensure alignment
if (!all(rownames(mapmycells_results) %in% colnames(rna.combined))) {
  stop("Cell names in MapMyCells results do not match Seurat object cell names.")
}

# Step 3: Merge MapMyCells annotations into Seurat metadata
rna.combined <- AddMetaData(
  object = rna.combined,
  metadata = mapmycells_results
)

# Inspect updated metadata
head(rna.combined@meta.data)



# Access the metadata
metadata <- rna.combined@meta.data

# Remove the first 4 digits followed by a space and the trailing "_<number>"
metadata$cluster_name2 <- gsub("^[0-9]{4} |_[0-9]+$", "", metadata$cluster_name)
# Save the updated metadata back into the Seurat object
rna.combined@meta.data <- metadata

# Inspect the updated metadata
head(rna.combined@meta.data)

# Remove the first 4 digits followed by a space and the trailing "_<number>"
metadata$subcluster_name2 <- gsub("^[0-9]{4} |_[0-9]+$", "", metadata$subcluster_name)
# Save the updated metadata back into the Seurat object
rna.combined@meta.data <- metadata

# Inspect the updated metadata
head(rna.combined@meta.data)


# Determine the number of clusters
n_clusters <- length(unique(rna.combined$subcluster_name2))

library(RColorBrewer)
# Get the "Paired" palette from RColorBrewer
palette <- brewer.pal(n = min(n_clusters, 12), name = "Accent")  # Adjust for max palette size

# Extend the palette if more than 12 clusters (optional)
if (n_clusters > 12) {
  palette <- colorRampPalette(brewer.pal(12, "Accent"))(n_clusters)
}

# Apply the palette to DimPlot
p <- DimPlot(
  object = rna.combined,
  group.by = "subcluster_name2",
  reduction = "umap_after_harmony_4",
  pt.size = 0.6,
  label = TRUE,
  label.size = 2
) + 
  scale_color_manual(values = palette)
plot_filename <- paste0(plot_directory, "/UMAP_ANNOTATED_pc42_res1_subcluster_name2_res1", ".pdf")
  #png(plot_filename, width = 1800, height = 600) 
  pdf(plot_filename, width = 11, height = 8)  # Adjust width and height as desired
  print(p)
  dev.off()
  
  p <- DimPlot(
  object = rna.combined,
  #group.by = "seurat_clusters",
  reduction = "umap_after_harmony_4",
  pt.size = 0.6,
  label = TRUE,
  label.size = 2
) 
plot_filename <- paste0(plot_directory, "/UMAP_ANNOTATED_pc42_res1_ann", ".pdf")
  #png(plot_filename, width = 1800, height = 600) 
  pdf(plot_filename, width = 11, height = 8)  # Adjust width and height as desired
  print(p)
  dev.off()
  
# Determine the number of clusters
n_clusters <- length(unique(rna.combined$subcluster_name2))

library(RColorBrewer)
# Get the "Paired" palette from RColorBrewer
palette <- brewer.pal(n = min(n_clusters, 12), name = "Accent")  # Adjust for max palette size

# Extend the palette if more than 12 clusters (optional)
if (n_clusters > 12) {
  palette <- colorRampPalette(brewer.pal(12, "Accent"))(n_clusters)
}

# Apply the palette to DimPlot
p <- DimPlot(
  object = rna.combined,
  group.by = "subcluster_name2",
  reduction = "umap_after_harmony_4",
  pt.size = 0.6,
  label = TRUE,
  label.size = 2
) + 
  scale_color_manual(values = palette) +
  theme(legend.position = "none") 
plot_filename <- paste0(plot_directory, "/UMAP_ANNOTATED_pc42_res1_subcluster_name2", ".pdf")
  #png(plot_filename, width = 1800, height = 600) 
  pdf(plot_filename, width = 11, height = 8)  # Adjust width and height as desired
  print(p)
  dev.off()
  
````
Plots
Top 5 markers for each cluster
Genes included only if they are expressed in atleast 25% of cells, hence not testing infrequently expressed genes.
adjusted p-value threshold used here is 0.05
genes ordered also by average log FC values
```{r}

 # Set a directory to save the plots
plot_directory <- "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering"


# Create the directory if it doesn't exist
dir.create(plot_directory, showWarnings = FALSE)

#delete old annotation
Idents(rna.combined) <- rna.combined@meta.data$seurat_clusters
new.cluster.ids<-c("dSPNs Mtx","iSPNs Mtx","iSPNs Mtx","eSPNS","Deep layer CT", "SPNs like","iSPNs Patch","Splatter", "dSPNs Patch", "SPNs like", "dSPNs Patch", "LAMP5-LHX6 Int", "CGE Int", "MGE Int", "dSPNs CHST9+","Splatter","iSPNs Patch HTR7+","SPNs like", "iSPNs Mtx","iSPNs Mtx", "dSPNs Mtx","iSPNs Mtx","LAMP5-LHX6 Int" )
names(new.cluster.ids) <- levels(rna.combined)
rna.combined<- RenameIdents(rna.combined, new.cluster.ids)
rna.combined$clusters <- Idents(rna.combined)   
saveRDS(rna.combined_all, file="Z:/dmclab/Marta/PD/snRNA human putamen/data/rna.combined_res1_annotated_2_clinicaldata_ann_allen.rds")

library(dplyr)
wnn_markers <- FindAllMarkers(rna.combined, only.pos = TRUE, min.pct = 0.05, logfc.threshold = 0.05, recorrect_umi = FALSE)
wnn_markers <- wnn_markers[wnn_markers$p_val_adj < 0.05, ]
write.csv(wnn_markers, file = "Z:/dmclab/Marta/PD/snRNA human putamen/Subclustering/Harmony_2/cluster_markers-combined_snRNA_subclustering_annotated_res1__allen.csv", col.names = TRUE, row.names = TRUE)
wnn_markers<- read.csv("Z:/dmclab/Marta/PD/snRNA human putamen/Subclustering/Harmony_2/cluster_markers-combined_snRNA_subclustering_annotated_harmony2.csv")
wnn_markers<- wnn_markers[!grepl("^AC|^AP|^AL|RIK$", wnn_markers$gene), ]
top5_2 <- wnn_markers %>%
  group_by(cluster) %>%
  slice_max(n = 5, order_by = avg_log2FC) %>%
  filter(row_number() %in% 1:5)

#add clinical data
library(readxl)
clinical_data<- read_xlsx("Z:/dmclab/Marta/PD/snRNA human putamen/clinical_data.xlsx")



# Extract the metadata from the Seurat object
rna_metadata <- rna.combined@meta.data

# Create a subset of clinical_data that includes only relevant columns
# Ensure clinical_data and rna_metadata have a common sample identifier
clinical_subset <- clinical_data[, c("SampleID", "Sex","Age", "Braak", "PMI")]



# Merge the clinical data with Seurat metadata based on 'SampleID'
# This will match rows where SampleID matches between the two data frames
merged_metadata <- merge(rna_metadata, clinical_subset, by = "SampleID", all.x = TRUE)

# Add the updated metadata back to the Seurat object
rna.combined <- AddMetaData(rna.combined, metadata = merged_metadata)
metadata <- rna.combined@meta.data

braak_colors <- c(
  "0" = "gray", 

  "3" = "#D9BFBF", 
  "4" = "#a34246ff", 
  "5" = "#79090fff", 
  "6" = "#330000"
)
rna.combined$Braak <- factor(rna.combined$Braak, levels = names(braak_colors))
  p2<-DimPlot(object = rna.combined, split.by ="Braak", reduction = "umap_after_harmony_4",pt.size = 0.4, label = F, label.size = 6, raster=FALSE,ncol=3) + ggtitle("Clusters in human snRNAseq dataset") +scale_color_manual(values = braak_colors)
plot_filename <- paste0(plot_directory, "/UMAP_sub_40PC_RES1_HD_splitby_braak_sample_g", ".pdf")
  #png(plot_filename, width = 1800, height = 600) 
  pdf(plot_filename,width=16, height=6)  # Adjust width and height as desired
  print(p2)
  dev.off()

  


cl.colors<- c("#e080b9","#4e6e63","#b6dddb","#a2cae9", "#f3c9d9","#bc7ed4","#878787","#0a4347ff","#d04177","#6c9b9c","#f78954","#a26074","#f69e92","#4b053c")

cl.colors<- c("#0a4347ff","#e080b9","#bc7ed4","#dace7eff","#99B9B5","#f3c9d9","#878787","#b6dddb","#102c52ff","#516673","#677991","#a2cae9","#f69e92" )


p<-DimPlot(object = rna.combined,  reduction = "umap_after_harmony_4",  pt.size = 0.8, label = F,cols=alpha(cl.colors,0.8) ,label.size = 4) + ggtitle("SubClusters for human snRNAseq dataset")
  plot_filename <- paste0(plot_directory, "/UMAP_42PC_RES1_annotated_cols_", ".pdf")
  pdf(plot_filename, width = 10, height = 9)  # Adjust width and height as desired
  print(p)
  dev.off()



my_table<- as.data.frame(table("cluster" = rna.combined@active.ident, rna.combined@meta.data$orig.ident))


p <- DimPlot(object = rna.combined, group.by = "seurat_clusters", reduction = "umap_after_harmony_4",
             pt.size = 0.6, label = TRUE, label.size = 6)+ggtitle("Clustering of all human samples snRNA seq data")

# Create the table_plot and use the cluster colors
table_plot <- ggplot(my_table, aes(x = cluster, y = Freq, fill = cluster)) +
  geom_col() +

  labs(x = "counts", y = "Cluster") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank())

plot<- p+ table_plot
#plot<- grid.arrange(p, table_plot, ncol = 2)
plot_filename <- paste0(plot_directory, "/ALL_UMAP_table_42PC_RES1", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
pdf(plot_filename, width = 16, height = 8)  # Adjust width and height as desired
print(plot)
dev.off()



table2<-as.data.frame(table("cluster" = rna.combined@active.ident, rna.combined@meta.data$Sample_group))
write.csv(table2, file="Z:/dmclab/Marta/PD/snRNA human putamen/Subclustering/Harmony_2/Nuclei_per_sample_group.csv")

#table2$color<- c("#B95F89", "#0a4347" ,"#BDCCE6", "#140224", "gray","#B95F89", "#0a4347" ,"#BDCCE6", "#140224", "gray" )
# Create the bar plot
# Calculate percentages
# Determine the number of clusters
rna.combined@meta.data$orig.ident <- "snRNA seq experiments"
table<- table("cluster" = rna.combined@active.ident, rna.combined@meta.data$orig.ident, rna.combined@meta.data$Sample_group)
table<- as.data.frame(table)

# Calculate percentages
Percentage <- (table$Freq / sum(table$Freq)) * 100
table1<- cbind(table[1],table[3],table[4],Percentage)

Sample_groups<- unique(table$Var3)
table_final<- data.frame()
for (i in Sample_groups){
  table_group<- subset(table, Var3== i)
  Percentage <- (table_group$Freq / sum(table_group$Freq)) * 100
  table1<- cbind(table_group[1],table_group[3],table_group[4],Percentage, cl.colors)
  table_final<- rbind(table1, table_final)

}

p<-ggplot(table_final, aes(x = Var3, y = Percentage, fill=cluster)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(x = "Sample Group", y = "Percentage") +
  scale_fill_manual(values = table_final$cl.colors) +  # Assign colors
  theme_classic() +  # Set plot theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
plot_filename <- paste0(plot_directory, "/proportion_overall_stacked", ".pdf")
  #png(plot_filename, width = 1800, height = 600) 
  pdf(plot_filename, width = 5, height = 5)  # Adjust width and height as desired
  print(p)
  dev.off()  
  

#specific markers 
markers<-c("MBP","TNC","CX3CR1","SLC1A3","SOX10","PDGFRA","GFAP","ALDH1L1","SLC17A7","SLC17A6" ,"SLC32A1","GAD2","SST","NPY","PVALB", "CHAT", "ACHE","SLC18A3","GPR6","PPP1R1B","RGS9" ,"GPR88","ADORA2A","DRD2","PENK","PDYN","TAC1","DRD1","COL11A1","SEMA5B","OPRM1","TH")
markers<-c("PPP1R1B", "SLC32A1","GPR6","GPR88","ADORA2A","DRD1","DRD2", "KCNIP1","STXBP6","RXFP1","CRYM","CNR1","PENK","PDYN","TAC1","LYPD1","OPRM1","KREMEN1", "ID4","EPHA4","SEMA5B","SST","NPY","CHAT", "ACHE","MBP")
markers<- c("PPP1R1B","GPR6","GPR88","CALB1","COL11A1", "KREMEN1","ADORA2A","DRD2","PENK","PDYN","TAC1","DRD1","CRYM","CNR1","KCNIP1","STXBP6","RXFP1","LYPD1","OPRM1","ID4","EPHA4","SEMA5B","SLC32A1","ACHE","SST","NPY","CHAT", "MBP","CRYM", "GPR155", "DLK1","PVALB")
markers<- c("ADORA2A","DRD2","PENK","PDYN","TAC1","DRD1","CRYM","KCNIP1","RXFP1","SEMA5B","COL11A1","ADARB2","ACHE","SST","NPY")
mrk_espn<- c("RXFP1", "COL11A1", "PCDH8","OTOF" ,"TSHZ1","ADARB2", "SEMA5B","CPNE4", "GRIK1","CACNG5","CNR1","NETO1","SORCS2","PTPRT","DRD1","DRD2","DRD3")
all_genes<- c(d1_markers,st_markers)
DefaultAssay(rna.combined)<- "SCT"
custom_order<- c("dSPNs Mtx","iSPNs Mtx","eSPNS","SPNs like","dSPNs Patch","dSPNs CHST9+","iSPNs Patch","iSPNs Patch HTR7+")
DefaultAssay(rna.combined_sub)<- "SCT"
d3<-DotPlot(rna.combined_sub, features =markers)+
  scale_colour_gradientn(colours =rev(c("white","gray", "#B34B50", "#5F070C","#240705") )%>% rev()) + labs(y = "cluster")+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=-0.3))+
  scale_x_discrete(position = "bottom") +
  scale_y_discrete(limits = rev(custom_order))+
  theme(axis.text = element_text(family = "sans", size = 7),
        axis.title = element_text(family = "sans", size = 10),
        legend.key.size = unit(5, units = "mm"),
        legend.text = element_text(family = "sans", size = 7),
        legend.title = element_text(family = "sans", size = 10))+
 scale_size(range = c(1,8))
plot_filename <- paste0(plot_directory, "/top_markers_annotated_res1_ONLYSPN", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
pdf(plot_filename,width=8, height = 5) # Adjust width and height as desired
print(d3)
dev.off()

DefaultAssay(rna.combined) <- "SCT"

cl.colors <-  c("#048BA8","#B95F89","#BDCCE6","gray")
cl.colors<- c("#B95F89", "#0a4347","#856fa7ff","#140224","#595959ff")
p<-DimPlot(object = rna.combined,  reduction = "umap_after_harmony_4", pt.size = 0.6, label = F, cols=cl.colors, split.by="Sample_group") + ggtitle("SubClusters for human snRNAseq dataset")
plot_filename <- paste0(plot_directory, "/UMAP_annotated_sub_42pc_res0.4_split_col2", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 14, height = 6)  # Adjust width and height as desired
  print(p)
  dev.off()

#violin plot of markers
lapply(unique(ALLMRK), function(gene){
                          
   myplot1 <- VlnPlot(object = rna.combined, features = gene, pt.size = 0,log=F,  cols = cl.colors)

  plot_filename <- paste0(plot_directory, "/selected_markers_violin_", gene, ".pdf")
  pdf(plot_filename, width = 6, height = 3)  # Adjust width and height as desired
  print(myplot1)
  dev.off()
})

#umap of each gene 
DefaultAssay(rna.combined_sub)<- "SCT"
lapply(unique(markers), function(gene){
  tryCatch({
    myplot <- FeaturePlot(rna.combined_sub, features = gene, reduction = "umap_after_harmony_4", slot = "data", pt.size=0.6,
                          label=FALSE, raster=FALSE) & 
                          theme(legend.position = "right") &
                          #scale_color_gradientn(colors = c("gray","#60795bff", "#354332ff"), limits=c(0,3))
                          scale_color_gradientn(colors = c("gray","#B34B50", "#5F070C","#240705"), limits=c(0,7))
      

    plot_filename <- paste0(plot_directory, "/marker_umap_HD_norm_fig_sized_", gene, ".pdf")
    pdf(plot_filename, width = 10, height = 9)  # Adjust width and height as desired
    print(myplot)
    dev.off()

  }, error = function(e) {
    message(paste("Error with gene:", gene, " - Skipping to next..."))
    # Continue with the next gene if there's an error
  })

})







```

```{r message=FALSE, warning=FALSE}
#other method to check the DE with MAST instead of wilcox

de_genes4<- list()
de_genes5<- list()
de_genes6<- list()
de_genes<- list()


rna.combined_all<-rna.combined
rna.combined <- subset(
  x = rna.combined,
  cells = WhichCells(rna.combined, idents = setdiff(Idents(rna.combined), c("Splatter", "LAMP5-LHX6 Int", "CGE Int", "MGE Int")))
)
use_clusters <- rna.combined@active.ident
all_clusters <- as.character(sort(levels(unique(use_clusters))))
DefaultAssay(rna.combined) <- "RNA"
rna.combined <- NormalizeData(rna.combined)

options(digits=15)
nuclei_count <-as.data.frame(table("cluster" = rna.combined@active.ident, rna.combined@meta.data$Sample_group,  rna.combined@meta.data$Braak))

#check the total number of cells in cluster and each sample

cl_names <- c()
for (cl in 1:length(all_clusters)) {
    print(all_clusters[cl])
    Ctrl_cells <- colnames(rna.combined)[rna.combined@active.ident %in% all_clusters[cl] & rna.combined$Sample_group %in% "Control"]
    treated_cells <- colnames(rna.combined)[rna.combined@active.ident %in% all_clusters[cl] & rna.combined$Sample_group %in% "PD"]

    if (length(Ctrl_cells) < 1 | length(treated_cells) < 1){
        print("test")
        next
    }
    
    length_ct<- length(Ctrl_cells)
    length_tr<- length(treated_cells)
    
    if (length_ct > length_tr){
      
    num_cells_to_select <- length(treated_cells)
    selected_ct_cells <- sample(Ctrl_cells,num_cells_to_select)
    # 
    de_genes[[cl]] <- FindMarkers(rna.combined[["RNA"]], slot = "data", cells.1 = treated_cells, cells.2 = selected_ct_cells, test.use = "MAST", latent.vars = "SampleID", min.pct = 0.05, logfc.threshold = 0.2,            features = NULL)
      
    }else{
    
      #Randomly select the same number of cells from shuffled bigger group so to have the same number of cells in the conditions
    num_cells_to_select <- length(Ctrl_cells)
    selected_tr_cells <- sample(treated_cells,num_cells_to_select)
    # 
    de_genes[[cl]] <- FindMarkers(rna.combined[["RNA"]], slot = "data", cells.1 = selected_tr_cells, cells.2 = Ctrl_cells, test.use = "MAST", latent.vars = "SampleID", min.pct = 0.05, logfc.threshold = 0.2,      features = NULL)
    }
    
    de_genes[[cl]]$Gene <- rownames(de_genes[[cl]])
    
    cl_names <- union(cl_names, all_clusters[cl])
}
names(de_genes) <- cl_names
library(openxlsx)
write.xlsx(de_genes, file = "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/PD_vs_wt_DEGs-subcluster_res1_allen.xlsx")
# Specify the file path of your Excel workbook
install.packages("readxl")
library(readxl)
excel_file <-"Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/PD_braak4_vs_wt_DEGs-subcluster_res1_allen.xlsx"
de_genes4<- read_xlsx(excel_file)
# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes4 <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes4[[sheet_name]] <- sheet_data
}

excel_file <-"Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/PD_braak5_vs_wt_DEGs-subcluster_res1_allen.xlsx"
de_genes5<- read_xlsx(excel_file)
# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes5 <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes5[[sheet_name]] <- sheet_data
}

excel_file <-"Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/PD_braak6_vs_wt_DEGs-subcluster_res1_allen.xlsx"
# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes6 <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes6[[sheet_name]] <- sheet_data
}

excel_file <-"Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/PD_vs_wt_DEGs-subcluster_res1_allen.xlsx"

# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes[[sheet_name]] <- sheet_data
}




de<- bind_rows(de_genes, .id = "cluster")
de <- de %>% filter(p_val_adj < 0.0001)
de_up <- de %>% filter(avg_log2FC >0.2)
de_down <- de %>% filter(avg_log2FC < -0.2)
de<-rbind(de_up, de_down)

clusters<- unique(rna.combined@active.ident)


nuclei_count <-as.data.frame(table("cluster" = rna.combined@active.ident, rna.combined@meta.data$Sample_group))

# nuclei_count<-nuclei_count  %>% filter(Var2 == "PD")
 de_1<-de %>% group_by(cluster) %>%
  summarise(count = n())
 nuclei_count<- subset(nuclei_count, cluster!="Deep layer CT")
de_merge <- merge(de_1, nuclei_count, by = "cluster")
de_merge <- transform(de_merge, Perc = count / Freq)
total_genes<- sum(de_merge$Freq)
de_merge<-transform(de_merge, norm = Freq / total_genes)

table_gene_count<- data.frame()
 for (clust in clusters){
   # subset<- subset(rna.combined, Sample_group== "PD")
   subset_cluster<- subset(rna.combined,  idents =clust)
   subset_cluster <- subset_cluster[rowSums(subset_cluster) > 0, ]
   gene_count<- length(rownames(subset_cluster))
   cluster<- clust
   gene_counts_df<- as.data.frame(gene_count, cluster)
   table_gene_count<- rbind(table_gene_count,gene_counts_df )
 }
#or

table_gene_count <- data.frame()

for (clust in clusters) {
   # Subset the Seurat object for the current cluster
   subset_cl <- subset(rna.combined, idents = clust)
   
   # Debugging: Print the current cluster being processed
   print(paste("Processing cluster:", clust))

   # Total number of rows (genes)
   total_rows <- length(rownames(subset_cl))
   
   # Count rows where the sum is equal to 0 (same logic as rowSums() == 0)
   zero_sum_rows <- sum(rowSums(subset_cl) == 0)
   
   # Subtract zero-sum rows to get rows with non-zero sums
   gene_count <- total_rows - zero_sum_rows
   
   # Create a data frame for the result
   gene_counts_df <- data.frame(cluster = clust, gene_count = gene_count)
   
   # Append the result to the final table
   table_gene_count <- rbind(table_gene_count, gene_counts_df)
}

# Check the final result
print(table_gene_count)

table_gene_count$Experiment<- "PD"
cluster<- rownames(table_gene_count)
table_gene_count<- cbind(cluster, table_gene_count)
colnames(table_gene_count)<- c("cluster","total_genes","Experiment")
table_gene_count <- subset(table_gene_count,  cluster!="Deep layer CT")
de_merge <- merge(de_1, table_gene_count, by = "cluster")
#de_merge<- de_merge[,-c(3,5)]
colnames(de_merge)<- c("cluster","count","total_genes")
de_merge<-transform(de_merge, norm = count / total_genes)
#de_6ohda_merge <- merge(de_6ohda_merge, table_colors, by = "cluster")

# Create the bar plot
p<- ggplot(de_merge, aes(x = reorder(cluster, norm), y = norm)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Clusters", y = "Significant changed genes/ total genes in the cluster (p value<0.0001)", title = "Genes affected per group") +
  ylim(0,0.1)+
  scale_fill_discrete(name = "Cluster")+
  theme(axis.text.x = element_text(angle = 45))+
    coord_flip()+
  theme_classic()
plot_filename <- paste0(plot_directory, "/barplot_de_norm_only_spns_allgenes", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 5, height = 4)  # Adjust width and height as desired
  print(p)
  dev.off()
  
  
  
  
#Braak stages 
#filter datasets 
de4<- bind_rows(de_genes4, .id = "cluster")
de4 <- de4 %>% filter(p_val_adj < 0.0001)
de4 <- subset(de4,cluster!="Deep layer CT")
de4_up<- de4 %>% filter(avg_log2FC > 0.2)
de4_down<- de4 %>% filter(avg_log2FC < -0.2)
de4<- rbind(de4_up,de4_down)


de6<- bind_rows(de_genes6, .id = "cluster")
de6 <- de6 %>% filter(p_val_adj < 0.0001)
de6<- subset(de6, cluster!="Deep layer CT")
de6_up<- de6 %>% filter(avg_log2FC > 0.2)
de6_down<- de6 %>% filter(avg_log2FC < -0.2)
de6<- rbind(de6_up,de6_down)

de5<- bind_rows(de_genes5, .id = "cluster")
de5 <- de5 %>% filter(p_val_adj < 0.0001)
de5<- subset(de5, cluster!="Deep layer CT")
de5_up<- de5 %>% filter(avg_log2FC > 0.2)
de5_down<- de5 %>% filter(avg_log2FC < -0.2)
de5<- rbind(de5_up,de5_down)

nuclei_count <-as.data.frame(table("cluster" = rna.combined@active.ident, rna.combined@meta.data$Braak))

# three clusters
table(rna.combined@active.ident)
cluster.set <- unique(rna.combined$nFeature_RNA)
# overall nFeature
sum(rowSums(rna.combined[['RNA']]@counts) != 0)


nuclei_count4<-nuclei_count  %>% filter(Var2 %in% c("0","4"))
 de4_1<-de4 %>% group_by(cluster) %>%
  summarise(count = n())


nuclei_count5<-nuclei_count   %>% filter(Var2 %in% c("0", "5"))
 de5_1<-de5 %>% group_by(cluster) %>%
  summarise(count = n())

nuclei_count6<-nuclei_count  %>% filter(Var2 %in% c("0","6"))
 de6_1<-de6 %>% group_by(cluster) %>%
  summarise(count = n())

clusters<- unique(Idents(rna.combined))
table_gene_count_4<- data.frame()
 for (clust in clusters){
   subset<- subset(rna.combined, Braak %in% c("0","4"))
   subset_cluster<- subset(subset,  idents =clust)
   subset_cluster <- subset_cluster[rowSums(subset_cluster) > 0, ]
   gene_count<- length(rownames(subset_cluster))
   cluster<- clust
   gene_counts_df<- as.data.frame(gene_count, cluster)
   table_gene_count_4<- rbind(table_gene_count_4,gene_counts_df )
 }

table_gene_count_5<- data.frame()
 for (clust in clusters){
   subset<- subset(rna.combined, Braak %in% c("0","5"))
   subset_cluster<- subset(subset,  idents =clust)
   subset_cluster <- subset_cluster[rowSums(subset_cluster) > 0, ]
   gene_count<- length(rownames(subset_cluster))
   cluster<- clust
   gene_counts_df<- as.data.frame(gene_count, cluster)
   table_gene_count_5<- rbind(table_gene_count_5,gene_counts_df )
   
 }

table_gene_count_6<- data.frame()
 for (clust in clusters){
   subset<- subset(rna.combined, Braak %in% c("0","6"))
   subset_cluster<- subset(subset,  idents =clust)
   subset_cluster <- subset_cluster[rowSums(subset_cluster) > 0, ]
   gene_count<- length(rownames(subset_cluster))
   cluster<- clust
   gene_counts_df<- as.data.frame(gene_count, cluster)
   table_gene_count_6<- rbind(table_gene_count_6,gene_counts_df )
   
 }


table_gene_count_4$Braak<- "4"
cluster<- rownames(table_gene_count_4)
table_gene_count_4<- cbind(cluster, table_gene_count_4)
colnames(table_gene_count_4)<- c("cluster","total_genes","Braak")

de4_merge <- merge(de4_1, table_gene_count_4, by = "cluster")
de4_merge<-transform(de4_merge, norm = count / total_genes)
#de_6ohda_merge <- merge(de_6ohda_merge, table_colors, by = "cluster")

table_gene_count_5$Braak<- "5"
cluster<- rownames(table_gene_count_5)
table_gene_count_5<- cbind(cluster, table_gene_count_5)
colnames(table_gene_count_5)<- c("cluster","total_genes","Braak")

de5_merge <- merge(de5_1, table_gene_count_5, by = "cluster")
de5_merge<-transform(de5_merge, norm = count / total_genes)
#de_6ohda_merge <- merge(de_6ohda_merge, table_colors, by = "cluster")

table_gene_count_6$Braak<- "6"
cluster<- rownames(table_gene_count_6)
table_gene_count_6<- cbind(cluster, table_gene_count_6)
colnames(table_gene_count_6)<- c("cluster","total_genes","Braak")

de6_merge <- merge(de6_1, table_gene_count_6, by = "cluster")
de6_merge<-transform(de6_merge, norm = count / total_genes)
#de_mp_merge <- merge(de_mp_merge, table_colors, by = "cluster")


total_genes<- rbind(de4_merge,de5_merge, de6_merge)


# Create the bar plot
p<- ggplot(total_genes, aes(x = reorder(cluster, norm), y = norm, fill = Braak)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Clusters", y = "Significat changed genes/ total genes in the cluster (p value<0.0001)", title = "Genes affected per group") +
  ylim(0,0.1)+
  scale_fill_discrete(name = "Cluster")+
   facet_grid(.~Braak)+
  theme(axis.text.x = element_text(angle = 45))+
    coord_flip()+
  theme_classic()
plot_filename <- paste0(plot_directory, "/barplot_de_all_Braakstages_norm_allgenes", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 6, height = 3)  # Adjust width and height as desired
  print(p)
  dev.off()
  
  
  
```

```{r}


dataset_braak<- list("All"=de,"Braak_4"=de4, "Braak_5"=de5, "Braak_6"=de6) 

for (i in names(dataset_braak)){

#clusters <- unique(Idents(rna.combined))
#clusters<- c("striatum D1","Striatum D2" )

#Volcano all genes
de<- dataset_braak[[i]]
de <- subset(de, cluster!="Deep layer CT") 
  keyvals <- ifelse(
    de$avg_log2FC < -0.2, '#31869bff',
      ifelse(de$avg_log2FC > 0.2, '#963634ff',
        'black'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == '#963634ff'] <- 'Upregulated'
  names(keyvals)[keyvals == '#31869bff'] <- 'Dowregulated'
  

  
  p <- EnhancedVolcano(de,
                       lab = NA, FCcutoff = 0.5,
                       pCutoff = 10e-6,
                       x = "avg_log2FC", y = "p_val_adj",pointSize = 2.0,
                        labSize = 4.0, colCustom = keyvals) +
    ylim(0,150)+
    xlim(-4,4)+
    labs(x= "avg log2 FC", y="p val adjusted",title = paste("Volcano plots 6OHDA:"))+
    theme(legend.position="none")+
    theme_classic()
   plot_filename <- paste0(plot_directory, "/volcano_human_all_braak_", i,  ".pdf")
  pdf(plot_filename, width = 8, height = 6)  # Adjust width and height as desired
  print(p)
  dev.off()
  
  x = list(de_up= de_up$Gene, de_down= de_down$Gene)
if (!require(devtools)) install.packages("devtools")
devtools::install_github("yanlinlin82/ggvenn")
library(ggvenn)
mycol<- c("#963634ff","#31869bff")
up <- ggvenn(
  x, 
  fill_color = mycol,auto_scale = TRUE,
  stroke_size = 0.5, set_name_size = 4
  )
# Add title after initializing the plot
up <- up + ggtitle("Affected genes human putamen")
plot_filename <- paste0(plot_directory, "/venn_diag_subclustering",i, ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 4, height = 4)  # Adjust width and height as desired
  print(up)
  dev.off()

}
  
#venn diagragm for common up and dowmregulated genes
library(VennDiagram)
mycol<- c("#963634ff","#31869bff")
x = list(de_up= de_up$Gene, de_down= de_down$Gene)
if (!require(devtools)) install.packages("devtools")
devtools::install_github("yanlinlin82/ggvenn")
library(ggvenn)
up <- ggvenn(
  x, 
  fill_color = mycol,auto_scale = TRUE,
  stroke_size = 0.5, set_name_size = 4
  )
# Add title after initializing the plot
up <- up + ggtitle("Affected genes human putamen")
plot_filename <- paste0(plot_directory, "/venn_diag_subclustering", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 4, height = 4)  # Adjust width and height as desired
  print(up)
  dev.off()
  
  
y=list(de_60ohda= de_60hda_down$Gene, de_mp= de_mp_down$Gene)
down<- ggvenn(
  y, 
  fill_color = mycol,
  stroke_size = 0.5, set_name_size = 4
  )
# Add title after initializing the plot
down <- down + ggtitle("downregulated genes")
plot_filename <- paste0(plot_directory, "/venn_downreg", ".pdf")
  #png(plot_filename, width = 600, height = 600) 
  pdf(plot_filename, width = 13, height = 6)  # Adjust width and height as desired
  print(down)
  dev.off()  

  
  
clusters <- clusters[!(clusters %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter"))]
for (i in names(dataset_braak)){

  de_genes<- dataset_braak[[i]]
  #de_genes <- subset(de_genes, cluster!="Deep layer CT") 
for (clust in clusters) {
  subset_data_human <- subset(de_genes,  cluster== clust)
   subset_data_human <- subset_data_human[subset_data_human$p_val_adj < 0.0001 & (subset_data_human$avg_log2FC < -0.2 | subset_data_human$avg_log2FC > 0.2), ]
   de <- subset(de, cluster!="Deep layer CT") 
if (length(subset_data_human)==0){
  next
}else{
genes<- unique(subset_data_human$Gene)
genes <- genes[!(grepl("^GM", genes) | grepl("RIK$", genes))]

  clust_filename <- gsub("/", "_", clust)

keyvals <- ifelse(
    subset_data_human$avg_log2FC < -0.2, '#31869bff',
      ifelse(subset_data_human$avg_log2FC > 0.2, '#963634ff',
        'black'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == '#963634ff'] <- 'Upregulated'
  names(keyvals)[keyvals == '#31869bff'] <- 'Dowregulated'

  p <- EnhancedVolcano(subset_data_human,
                       lab = NA, FCcutoff = 0.5,
                       pCutoff = 10e-6,
                       x = "avg_log2FC", y = "p_val_adj",pointSize = 2.0,
                        labSize = 4.0, colCustom = keyvals) +
        ylim(0,150)+
    xlim(-4,4)+
    labs(x= "avg log2 FC", y="p val adjusted",title = paste("Combined Volcano plots for cluster IN human subclusters: ", clust))+
    theme_classic()

  plot_filename <- paste0(plot_directory, "/volcano_human_", clust_filename,"_",i, ".pdf")
  pdf(plot_filename, width = 8, height = 6)  # Adjust width and height as desired
  print(p)
  dev.off()
 }
}
}
clusters<- unique(de$cluster)
for (i in names(dataset_braak)){
  df1<- dataset_braak[[i]]
    df1<- subset(df1, cluster!= "Deep layer CT")
# for (clust in clusters){
# df1<- subset(df, cluster== clust)
# #   
# de<- subset(de, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))

  m1_forplots<- df1[!grepl("^GM", df1$Gene) & !grepl("RIK$", df1$Gene)  & !grepl("PTGDS", df1$Gene) & !grepl("^AC", df1$Gene)& !grepl("^AL", df1$Gene)&!grepl("^LINC", df1$Gene),]


most_negative <- m1_forplots[order(unique(m1_forplots$avg_log2FC)), ][1:30, ]
most_positive <- m1_forplots[order(unique(-m1_forplots$avg_log2FC)), ][1:15, ]

# Combine the results
result <- rbind(most_negative, most_positive)
result<- na.omit(result)

# Calculate the number of elements in each cluster
  cluster_counts <- result %>%
    count(cluster) %>%
    arrange(desc(n))
  
  # Reorder clusters based on the number of elements
  result <- result %>%
    mutate(cluster = factor(cluster, levels = cluster_counts$cluster))

plot<-ggplot(result, aes( cluster, reorder(Gene, -avg_log2FC),fill= avg_log2FC))+
       geom_tile()+
        scale_fill_gradient2(
    low = '#31869bff', 
    mid = 'white',
    high = '#963634ff', 
    midpoint = 0,limits = c(-2, 2), oob = scales::squish 
  ) +
  #facet_wrap(.~cluster, scales = "fixed", drop=TRUE)+
       theme_classic()+
  theme(
      axis.text.x = element_text( angle=90, vjust = 0.5, hjust = 0, size=10),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_blank()
    )+
  coord_flip()+
 ggtitle(paste("Most significantly changed genes")) 
 plot_filename <- paste0(plot_directory, "/Most_sign_genes_heat_lim2_",i,".pdf")
  pdf(plot_filename, width = 14, height = 4)   # Adjust width and height as desired
  print(plot)
  dev.off()  

  result$color_category <- ifelse(result$avg_log2FC > 0, "above_zero", "below_zero")

plot1 <- ggplot(result, aes(avg_log2FC, reorder(Gene, -avg_log2FC), color = color_category)) +
  geom_point(size = 4) +
  scale_color_manual(values = c("below_zero" = '#31869bff', "above_zero" = '#963634ff')) +
  scale_x_continuous(breaks = seq(-3, 3, by = 1), limits = c(-3, 3)) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0, size = 10),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_blank()
  ) +
  ggtitle(paste("Most significantly changed genes "))
    
  plot_filename <- paste0(plot_directory, "/Most_sign_genes_dot_nothres_",i,".pdf")
  pdf(plot_filename, width = 4, height = 8)   # Adjust width and height as desired
  print(plot1)
  dev.off()  
}


  
  
  
  
dataset_list<- list("Braak 4"=de4, "Braak 6"=de6)
clusters<- unique(de$cluster)

for (i in names(dataset_braak)) {
df<- dataset_braak[[i]]
for (clust in clusters){
 df1<- subset(df, cluster== clust)

  m1_forplots<- df1[!grepl("^GM", df1$Gene) & !grepl("RIK$", df1$Gene)  & !grepl("PTGDS", df1$Gene) & !grepl("^AC", df1$Gene)& !grepl("^AL", df1$Gene) &!grepl("^AC", df1$Gene)&!grepl("^LINC", df1$Gene),]

  most_negative <- m1_forplots[order(unique(m1_forplots$avg_log2FC)), ][1:15, ]
   most_positive <- m1_forplots[order(unique(-m1_forplots$avg_log2FC)), ][1:15, ]

  # Combine the results
  result <- rbind(most_negative, most_positive)
  result<- na.omit(result)

# # Calculate the number of elements in each cluster
#   cluster_counts <- result %>%
#     count(cluster) %>%
#     arrange(desc(n))
#   
#   # Reorder clusters based on the number of elements
#   result <- result %>%
#     mutate(cluster = factor(cluster, levels = cluster_counts$cluster))

  plot<-ggplot(result, aes(reorder(Gene, -avg_log2FC),avg_log2FC, fill= avg_log2FC))+
       geom_bar(stat="identity", width=0.7)+
        scale_fill_gradient2(
    low = '#31869bff', 
    mid = 'white',
    high = '#963634ff', 
    midpoint = 0,limits = c(-2, 2), oob = scales::squish 
  ) +
  ylim(-3,3)+
  #facet_wrap(.~cluster, scales = "fixed", drop=TRUE)+
       theme_classic()+
  theme(
      axis.text.x = element_text( angle=0, vjust = 0.5, hjust = 0, size=10),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_blank()
    )+
  coord_flip()+
 ggtitle(paste("Most significantly changed genes")) 
    clust<-  gsub("/", "_", clust)
 plot_filename <- paste0(plot_directory, "/Most_sign_genes_bar_","_",i,"_",clust, ".pdf")
  pdf(plot_filename, width = 5, height = 4)   # Adjust width and height as desired
  print(plot)
  dev.off() 
  
  
  plot<-ggplot(result, aes(cluster, reorder(Gene, -avg_log2FC),fill= avg_log2FC))+
       geom_tile()+
        scale_fill_gradient2(
    low = '#31869bff', 
    mid = 'white',
    high = '#963634ff', 
    midpoint = 0,limits = c(-4, 4), oob = scales::squish 
  ) +
  #facet_grid(experiment~., scales = "fixed", drop=TRUE)+
       theme_classic()+
  theme(
      axis.text.x = element_text( angle=90, vjust = 0.5, hjust = 0, size=10),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_blank()
    )+
  coord_flip()+
 ggtitle(paste0("Most significantly affected genes in", i)) 
  clust<-  gsub("/", "_", clust)
  plot_filename <- paste0(plot_directory, "/Most_sign_genes", "_",i,"_",clust, ".pdf")
  pdf(plot_filename, width = 8, height = 4)  # Adjust width and height as desired
  print(plot)
  dev.off()  
  }
} 

```

```{r}
library(ggplot2)
library(gplots)  # For reorder_within function
library(dplyr)
library(tidytext)  
library(progress)
library(writexl)
library(openxlsx)
excel_file <- "Z:/dmclab/Marta/Combined_snRNAseq/subclustering/new_allen_analysis/Combined-6OHDA_wt_DEGs-cluster_res15_nospl.xlsx"
de_genes_6ohda<- read_xlsx(excel_file)
# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes_6ohda <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes_6ohda[[sheet_name]] <- sheet_data
}

de_genes_6ohda<- bind_rows(de_genes_6ohda, .id = "cluster")
de_genes_6ohda <- de_genes_6ohda %>% filter(p_val_adj < 0.0001)
de_genes_6ohda_up <- de_genes_6ohda %>% filter(avg_log2FC > 0.2)
de_genes_6ohda_down <- de_genes_6ohda %>% filter(avg_log2FC < -0.2)
de_genes_6ohda<- rbind(de_genes_6ohda_up,de_genes_6ohda_down)
excel_file <- "Z:/dmclab/Marta/Combined_snRNAseq/subclustering/new_allen_analysis/Combined-mp15_18_wt_DEGs-cluster_res15_nospl.xlsx"
de_genes_mp<- read_xlsx(excel_file)
# Get the names of all sheets in the workbook
sheet_names <- excel_sheets(excel_file)

# Create an empty list to store data frames
de_genes_mp <- list()

# Loop through each sheet and read it into a data frame
for (sheet_name in sheet_names) {
  sheet_data <- read_excel(excel_file, sheet = sheet_name)
  de_genes_mp[[sheet_name]] <- sheet_data
}

de_genes_mp<- bind_rows(de_genes_mp, .id = "cluster")
de_genes_mp <- de_genes_mp %>% filter(p_val_adj < 0.0001)
de_genes_mp_up <- de_genes_mp %>% filter(avg_log2FC > 0.2)
de_genes_mp_down <- de_genes_mp %>% filter(avg_log2FC < -0.2)
de_genes_mp<- rbind(de_genes_mp_up,de_genes_mp_down)

excel_file <- "Z:/dmclab/Marta/Combined_snRNAseq/subclustering/Shared_genes_subluctering.xlsx"
de_genes_cm<- read_xlsx(excel_file)
de_genes_cm<- subset(de_genes_cm, agreement== "Convergent")
# Get the names of all sheets in the workbook
de5<- subset(de5, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))
de4<- subset(de4, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))
de6<- subset(de6, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))
de<- subset(de, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))

de_genes_6ohda<- subset(de_genes_6ohda, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))

de_genes_mp<- subset(de_genes_mp, !(cluster %in% c("Deep layer CT", "LAMP5-LHX6 Int", "MGE Int", "CGE Int", "Splatter")))
m3 <- data.frame(matrix(ncol = 6, nrow = 0))

#(substr(de[i, 1], 1, 2) == substr(de_genes_6ohda[j, 1], 1, 2) &&
pb <- progress_bar$new(total = nrow(de4))
for (i in 1:nrow(de4)){
  #   pb$tick()
  # Sys.sleep(1/nrow(de))
  for (j in 1:nrow(de_genes_6ohda)){
  if  (toupper(de4[i, 7]) == toupper(de_genes_6ohda[j, 7])){
      m3[nrow(m3)+1, 1] <- de4[i, 1]
      m3[nrow(m3),2] <- de_genes_6ohda[j, 1]
      m3[nrow(m3), 3] <- de4[i, 7]
      m3[nrow(m3), 4] <- de4[i, 3]
      m3[nrow(m3), 5] <- de_genes_6ohda[j, 3]
       if (m3[nrow(m3), 4]*m3[nrow(m3), 5] < 0){
        m3[nrow(m3), 6] <- "Divergent"}
       else {
        m3[nrow(m3), 6] <- "Convergent"}
      }
    }
}
colnames(m3)<- c("Cluster_human","Cluster_6ohda","Gene","avg_log2FC_human","avg_log2FC_60hda", "agreement" )

library(writexl)

write.xlsx(m3, file = "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/common_genes_PD_braak4_6ohda_vs_wt_DEGs_p0001_latest_noclust.xlsx")


m3<- read_xlsx("Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/common_genes_PD_6ohda_vs_wt_DEGs_p0001_latest_noclust.xlsx")
m3_filter <- subset(m3, agreement == "Convergent")
mean<- cbind(m3_filter$avg_log2FC_6ohda, m3_filter$avg_log2FC_human)
m3_filter$mean <- rowMeans(mean)
# Sort correctly without using `unique()`
most_negative <- m3_filter[order(m3_filter$mean), ][1:30, ]  # Smallest 30 means
most_positive <- m3_filter[order(-m3_filter$mean), ][1:40, ]

# Combine the results
result <- rbind(most_negative, most_positive)
result<- na.omit(result)


result<- result[,-c(6,7)]
result1<- result[,-c(2,4)]
result1$experiment <- "Human"
colnames(result1)<- c("cluster","gene","avg_log2FC","experiment")
result2<- result[,-c(1,5)]
result2$experiment <- "6OHDA"
colnames(result2)<- c("cluster","gene","avg_log2FC","experiment" )

res_long <- rbind(result1, result2)
plot<-ggplot(res_long, aes( cluster, reorder(gene, -avg_log2FC),fill= avg_log2FC))+
       geom_tile()+
        scale_fill_gradient2(
    low = '#31869bff', 
    mid = 'white',
    high = '#963634ff', 
    midpoint = 0,limits = c(-1, 1), oob = scales::squish 
  ) +
  facet_grid(experiment~., scales = "fixed", drop=TRUE)+
       theme_classic()+
  theme(
      axis.text.x = element_text( angle=90, vjust = 0.5, hjust = 0, size=10),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_blank()
    )+
  coord_flip()+
 ggtitle(paste("Most significantly commonly genes with 6ohda")) 
  plot_filename <- paste0(plot_directory, "/Most_sign_shared_genes_human_6ohda_lim1_5_braak4", ".pdf")
  pdf(plot_filename, width = 8, height = 4)  # Adjust width and height as desired
  print(plot)
  dev.off()  


de_genes_mp<-subset(de_genes_mp, subset= cluster != "Unclassified")
m4 <- data.frame(matrix(ncol = 6, nrow = 0))
#substr(de[i, 1], 1, 2) == substr(de_genes_mp[j, 1], 1, 2) &&
pb <- progress_bar$new(total = nrow(de4))
for (i in 1:nrow(de4)){
  #   pb$tick()
  # Sys.sleep(1/nrow(de))
  for (j in 1:nrow(de_genes_mp)){
  if  (toupper(de[i, 7]) == toupper(de_genes_mp[j, 7])){
      m4[nrow(m4)+1, 1] <- de4[i, 1]
      m4[nrow(m4),2] <- de_genes_mp[j, 1]
      m4[nrow(m4), 3] <- de4[i, 7]
      m4[nrow(m4), 4] <- de4[i, 3]
      m4[nrow(m4), 5] <- de_genes_mp[j, 3]
       if (m4[nrow(m4), 4]*m4[nrow(m4), 5] < 0){
        m4[nrow(m4), 6] <- "Divergent"}
       else {
        m4[nrow(m4), 6] <- "Convergent"}
      }
    }
}
colnames(m4)<- c("Cluster_human","Cluster_mp","Gene","avg_log2FC_human","avg_log2FC_mp", "agreement" )
write.xlsx(m4, file = "Z:/dmclab/Marta/PD/snRNA human putamen/new_allen_clustering/subclustering/common_genes_Braak4_mp_vs_wt_DEGs_p0001_latest_noclust_harmony2.xlsx")



m4<- read_xlsx("Z:/dmclab/Marta/PD/snRNA human putamen/Subclustering/Harmony_2/common_genes_braak4_mp_vs_wt_DEGs_p0001_latest_noclust_harmony2.xlsx")

m4_filter <- subset(m4, agreement == "Convergent")
mean<- cbind(m4_filter$avg_log2FC_mp, m4_filter$avg_log2FC_human)
m4_filter$mean <- rowMeans(mean)
most_negative <- m4_filter[order(m4_filter$mean), ][1:40, ]
most_positive <- m4_filter[order(-m4_filter$mean), ][1:50, ]

# Combine the results
result <- rbind(most_negative, most_positive)
result<- na.omit(result)


result<- result[,-c(6,7)]
result1<- result[,-c(2,4)]
result1$experiment <- "Human"
colnames(result1)<- c("cluster","gene","avg_log2FC","experiment")
result2<- result[,-c(1,5)]
result2$experiment <- "Mitopark"
colnames(result2)<- c("cluster","gene","avg_log2FC","experiment" )

res_long <- rbind(result1, result2)
plot<-ggplot(res_long, aes( cluster, reorder(gene, -avg_log2FC),fill= avg_log2FC))+
       geom_tile()+
        scale_fill_gradient2(
    low = '#31869bff', 
    mid = 'white',
    high = '#963634ff', 
    midpoint = 0,limits = c(-1,1 ), oob = scales::squish 
  ) +
  facet_grid(experiment~., scales = "fixed", drop=FALSE)+
       theme_classic()+
  theme(
      axis.text.x = element_text( angle=90, vjust = 0.5, hjust = 0, size=10),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_blank()
    )+
  coord_flip()+
 ggtitle(paste("Most significantly commonly genes with mitpark")) 
  plot_filename <- paste0(plot_directory, "/Most_sign_shared_genes_human_braak4_mp_lim1_5", ".pdf")
  pdf(plot_filename, width = 8, height = 4)  # Adjust width and height as desired
  print(plot)
  dev.off()  





m5 <- data.frame(matrix(ncol = 8, nrow = 0))

#(substr(de[i, 1], 1, 2) == substr(de_genes_6ohda[j, 1], 1, 2) &&
pb <- progress_bar$new(total = nrow(de))
for (i in 1:nrow(de)){
  #   pb$tick()
  # Sys.sleep(1/nrow(de))
  for (j in 1:nrow(de_genes_cm)){
  if  (toupper(de[i, 7]) == toupper(de_genes_cm[j, 2])){
      m5[nrow(m5)+1, 1] <- de[i, 1]
      m5[nrow(m5),2] <- de_genes_cm[j, 1]
      m5[nrow(m5), 3] <- de[i, 7]
      m5[nrow(m5), 4] <- de[i, 3]
      m5[nrow(m5), 5] <- de_genes_cm[j, 3]
      m5[nrow(m5), 6] <- de_genes_cm[j, 4]
       if (m5[nrow(m5), 4]*m5[nrow(m5), 5] < 0){
        m5[nrow(m5), 7] <- "Divergent_mp"}
       else {
        m5[nrow(m5), 7] <- "Convergent_mp"}
       if (m5[nrow(m5), 4]*m5[nrow(m5), 6] < 0){
        m5[nrow(m5), 8] <- "Divergent_6ohda"}
       else {
        m5[nrow(m5), 8] <- "Convergent_6ohda"}      
      
      }
    }
}



colnames(m5)<- c("Cluster_human","Cluster_mouse","Gene","avg_log2FC_human","avg_log2FC_mp","avg_log2FC_6ohda", "agreement_mp" , "agreement_6ohda" )


library(writexl)

write.xlsx(m5, file = "Z:/dmclab/Marta/PD/snRNA human putamen/Subclustering/Harmony_2/common_genes_with_mouse_DEGs_p0001_latest_noclust_harmony2_noint.xlsx")


m5_filter <- subset(m5, agreement_mp == "Convergent_mp")
m5_filter <- subset(m5_filter, agreement_6ohda == "Convergent_6ohda")


result<- m5_filter[,-c(7,8)]
result1<- result[,-c(2,5,6)]
result1$experiment <- "Human"
colnames(result1)<- c("cluster","gene","avg_log2FC","experiment")
result2<- result[,-c(1,4,5)]
result2$experiment <- "6OHDA"
colnames(result2)<- c("cluster","gene","avg_log2FC","experiment" )
result3<- result[,-c(1,4,6)]
result3$experiment <- "Mitopark"
colnames(result3)<- c("cluster","gene","avg_log2FC","experiment" )

res_long <- rbind(result1, result2, result3)
library(ggplot2)

plot <- ggplot(res_long, aes(cluster, reorder(gene, -avg_log2FC), fill = avg_log2FC)) +
  geom_tile() +
  geom_text(aes(label = round(avg_log2FC, 2)), size = 3) +  # Add text labels with rounded avg_log2FC values
  scale_fill_gradient2(
    low = '#31869bff',
    mid = 'white',
    high = '#963634ff',
    midpoint = 0,
    limits = c(-1, 1),
    oob = scales::squish
  ) +
  facet_grid(experiment ~ ., scales = "fixed", drop = TRUE) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0, size = 10),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_blank()
  ) +
  coord_flip() +
  ggtitle("Most significantly commonly genes with mouse models")

plot_filename <- paste0(plot_directory, "/Most_sign_shared_genes_human_mouse_noclust_aharmony2_flip3", ".pdf")
pdf(plot_filename, width = 8, height = 4)  # Adjust width and height as desired
print(plot)
dev.off()


```
